---
title: "m4&5"
author: "Andrew Estes"
date: '2022-11-17'
output: pdf_document
---

```{r, echo=FALSE, warning=FALSE, error=FALSE,message=FALSE}
library(readxl)
library(tidyverse)
library(dplyr)
library(lubridate)
library(colorspace)
library(patchwork)
library(forecast)
library(zoo)
library(xts)
library(janitor)
library(knitr)
library(TTR)
library(viridis)
library(randtests)
library(gridExtra)
library(plotly)
```

```{r, echo=FALSE, warning=FALSE, error=FALSE,message=FALSE}
orig.nba <- read.csv("nba2018.csv")
nba <- orig.nba
```

# 1) Allows the user to choose two teams that will be compared by the graph(s) in the app.

# 2) Allows the user to choose two variables that will be compared in a scatter plot. 

Note: to do this, you might investigate the .data pronoun in ggplot, 
which can be used like this: 
  aes(.data[[x_var]], .data[[y_var]]), 
where x_var and y_var represent the names of variables stored as character strings.

# 3) Creates a scatter plot of the two chosen variables, using color to show which team each data point belongs to.

# 4) Uses plotly to render the graph in the app, 
either by creating the graph with ggplot, then using ggplotly to create the plotly graph, OR 
by using plot_ly and related commands directly. 
You'll need to use the renderPlotly and plotlyOutput commands to do that.

# 5) Creates nicely-formated pop-up text for each data point that contains 
each player's name, along with any other information you believe is valuable.
```{r}
#scatterplot Offensive Rebounds (OREB) vs Defensive Rebounds (DREB)
p <- ggplot(nba) +
  geom_point(aes(x=DREB, y=OREB, color=TEAM)) 
ggplotly(p)

```

