---
title: "Using manipulate for Interactive Graphs"
author: "Scott Thatcher"
date: "11/5/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(manipulate)
```


## Introduction to Interactive Visualizations

Interactive visualizations can...

- map additional variables (e.g. time as time!),

- allow slice-by-slice exploration,

- adjust aesthetics,

- zoom in and out,

- set up, then trigger a computation, etc., and

- provide pop-up feedback to the user.

[Visit Gapminder](https://www.gapminder.org/tools/), for a refresher!

## `manipulate` creates simple interactive plots in RStudio.

```{r eval=FALSE}
manipulate(<plot code>, <control1>, <control2>, ...)
```

- `<plot code>` Code that generates a plot. Multiple commands can be put inside of {} .

- `<control>` Expression of the form `param = control(…)`

  - `slider` Choose a numeric value from a continuum.

  - `picker` Choose a value from a list.
  
  - `checkbox` Toggle true/false values.
  
  - `button` Rerun `<plotcode>` with a parameter set to `TRUE`.
  
```{r}
# Here's an example:
manipulate(hist(rchisq(1000, df=x)), x=slider(1, 40))
```

```{r}
# Let's keep the axis constant and show df.
# Note the use of {}'s for multiple commands.
manipulate({
  hist(rchisq(1000, df=x), xlim=c(0, 80))
  text(x=70, y=150, label=paste0("df = ", x))
  },
  x=slider(1, 40)
)
```


```{r}
# Here's a fancier example:
manipulate({
  if(gg==TRUE){
    ggplot(data.frame(x=rchisq(1000, df=x))) +
      geom_histogram(aes(x=x), breaks=seq(0, 80 , 5))
  } else {
    hist(rchisq(1000, df=x), xlim=c(0, 80))
  }},
  x=slider(1, 40),
  gg=button(label="GGPlot!")
)
```

## Making an Interactive Visualization

- Step 1: Plan it out!

- Step 2: Make a static graph.

- Step 3: Replace hard-coded parameters with adjustable values.

## You try!

Use manipulate to visualize a statistical principle:

- The median is resistant to outliers, but the mean is not.

- A point separated horizontally from the rest of the data exerts
  a large influence on the regression line.
  
- The shape of a normal distribution depends only on mean and
  standard deviation.

- Varying binwidth in a histogram changes the balance between
  detail and summary.

## Details to watch for.

- When you animate a graph, do the axes stay constant, or does the
  graph "jump around"?
  
  - Changes caused by user intervention are easier to see if
    axes/other graph elements stay constant.
  
  - You might need to override R's default axis scaling, etc.
  
- Are control ranges appropriate?

- Are controls intuitive/well-explained?

- What data to use?

  - Fixed or random?
  
  - Simulated or real?
