---
title: "lab2b_estes"
author: "Andrew Estes"
date: "9/5/2021"
output: word_document
---

#### Installing Packages and Data Set
```{r}
library(tidyr)
library(tidyverse)
kim <- read.csv("C:\\Users\\andre\\OneDrive\\Desktop\\PDAT 610\\Module2b\\Clean-KimData.csv")
head(kim)
```


## Question 1: Select "male" OR "first year" and compare to "male AND first year" 
#### Answer: there were 229 observations of "male" OR "first year" compared to 43 observations of "male AND first year" 
```{r}
maleORfirst <- filter(kim, Semester < 2 | Gender == "M" )
count(maleORfirst)
```

\newpage
## Question 2: Calculate BMI 
```{r}
#adding numerical BMI to the dataframe
kimBMI <- mutate(kim, BMI = 703 *(Weight/(h=Height^2)))

#adding categorical TRUE/FALSE determining status of obese to dataframe
kimData <- mutate(kimBMI, Obese = BMI > 30)
head(kimData)
```

\newpage
## Question 3: Calculate BMI for each year
```{r}
#dividing the semeesters into 2 to calculate the year
#adding 1 to the calculation so Freshman are Year 1 instead of Year 0
kimData2 <- kimData %>% 
  mutate(Year = round(Semester/2)+1)
head(kimData2)

#grouping the obesity rates by Year
df <- kimData2 %>% group_by(Year)
df %>% summarise (
          avg = mean(BMI, na.rm = TRUE)
)
```

\newpage
## Question 4: Calculate the obesity percentage for each year
```{r error = TRUE}
#dr kim's answer
KimData <-kimData2 %>% 
  mutate(BMI=703*(Weight/Height^2))%>%
  mutate(Obese=BMI>=30) %>%
  mutate(Year = round(Semester/2))%>%
  group_by(Year) %>% 
  
  #filtering obesity by year
firstYearObese <- filter(df, Year == 1 & BMI > 30)
secondYearObese <- filter(df, Year == 2 & BMI > 30)
thirdYearObese <- filter(df, Year == 3 & BMI > 30)
fourthYearObese <- filter(df, Year == 4 & BMI > 30)
fifthYearObese <- filter(df, Year == 5 & BMI > 30)
sixthYearObese <- filter(df, Year == 6 & BMI > 30)


  #filtering data by year
firstYear <- filter(df, Year == 1)
secondYear <- filter(df, Year == 2)
thirdYear <- filter(df, Year == 3)
fourthYear <- filter(df, Year == 4)
fifthYear <- filter(df, Year == 5)
sixthYear <- filter(df, Year == 6)

  #dividing total number of obesity/yr by total respondents/yr
obeseFirstYr <- count(firstYearObese)/count(firstYear)
obeseSecondYr <- count(secondYearObese)/count(secondYear)
obeseThirdYr <- count(thirdYearObese)/count(thirdYear)
obeseFourthYr <- count(fourthYearObese)/count(fourthYear)
obeseFifthYr <- count(fifthYearObese)/count(fifthYear)
obeseSixthYr <- count(sixthYearObese)%/%count(sixthYear)

  #output of the above division
obeseFirstYr 
obeseSecondYr
obeseThirdYr
obeseFourthYr
obeseFifthYr
obeseSixthYr
```

\newpage
## Question 5: Analyze feet size by gender
#### Answer: I removed all observations over 2 standard deviations away from the mean
#### The male average drops from an initial value of 11.51 to 10.81
#### The male sd drops from an initial value 8.19 to 1.31
  
#### The female average drops from an initial value of 7.99 to 7.96
#### The female SD drops from an initial value 1.33 to 1.17
```{r}
#male analysis
maleDF <- filter(df, Gender == "M")
mean(maleDF$Shoe.Size, na.rm = TRUE)
sd(maleDF$Shoe.Size, na.rm = TRUE)

maleDF2 <- filter(maleDF, Shoe.Size < 20)
mean(maleDF2$Shoe.Size, na.rm = TRUE)
sd(maleDF2$Shoe.Size, na.rm = TRUE)

maleDF3 <- filter(maleDF2, Shoe.Size > 7.86 & Shoe.Size < 13.9)
mean(maleDF3$Shoe.Size, na.rm = TRUE)
sd(maleDF3$Shoe.Size, na.rm = TRUE)


#female analysis
femaleDF <- filter(df, Gender == "F")
mean(femaleDF$Shoe.Size, na.rm = TRUE)
sd(femaleDF$Shoe.Size, na.rm = TRUE)

femaleDF2 <- filter(femaleDF, Shoe.Size > 5.33 & Shoe.Size < 10.65)
mean(femaleDF2$Shoe.Size, na.rm = TRUE)
sd(femaleDF2$Shoe.Size, na.rm = TRUE)
```




