---
title: "lab3_estes"
author: "andrew.estes"
date: "9/11/2021"
output: pdf_document
---

Install Data and Packages
```{r}
#install.packages("AmesHousing")
#tinytex::install_tinytex()
library(AmesHousing)
library(tidyverse)
library(tinytex)
AmesData <- make_ames()
#view(AmesData)
#str(AmesData)
```

Create a new column in the AmesData data frame which has a value of 1 if the house is built in year 2000 or later or 0 if it is built before year 2000 using the following method:
1) IF-ELSE
2) FOR
3) SAPPLY
4) Question 3 - Question 1

\newpage
## 1) ifelse(name this column – w3ifelse). What is the first 10 elements of w3ifelse?
```{r}
AmesData2 <- AmesData %>% 
  select(Year_Built) %>%
  mutate(w3ifelse = ifelse(Year_Built > 1999, 1, 0))

head(AmesData2, 10)
view(AmesData2)
```

\newpage
## 2) if-Else and a For loop (name this column – w3for). What is the first 10 elements of w3for?
```{r}
AmesData3 <- AmesData2


  w3for = for (i in 1:nrow(AmesData3)){
    if(AmesData3$Year_Built[i] < 2000){
        AmesData3$Year_Built[i] <- 0
    } else {
        AmesData3$Year_Built[i] <- 1
    }
  }

head(AmesData3, 10)
```

\newpage
## 3) Build your own function and use sapply (name this column - w3apply). What is the first 10 elements of w3apply?
```{r}
woo <- function(x){
  if(is.na(x)) return(x)
  else if (x <= 1999) return (0)
  else return (1)
}

AmesData4 <- AmesData2 %>% 
  select(Year_Built, w3ifelse) %>%
  mutate(w3apply = sapply(AmesData2$Year_Built, woo))
         
head(AmesData4)

```

\newpage
## 4)	Create a column w3diff which is the difference of the two columns w3ifelse and w3apply. 
DO NOT use a for loop or the apply set of functions for this step. Simple subtraction will work. 
What is the total sum of this column? 
```{r}
AmesDataFinal <- AmesData4 %>%
  mutate(w3diff = w3apply - w3ifelse)

head(AmesDataFinal, 10)
```

