---
title: "Module 2A"
author: "Andrew Estes"
date: "8/22/2021"
output: word_document
---

### Pre-Analysis Set-Up
##### Installing Packages
```{r, warning = FALSE, message = FALSE, results='hide'}
library(ggpubr)
library(car)
library(caret)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(MASS)
library(sensemakr)
```


##### Importing and View Data
```{r}
abalone.url <-
  "https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data"

abalone.data <- read.csv(abalone.url, header=FALSE) #data does not have variable names

head(abalone.data)
```

\newpage
### 1) What is the class of the data abalone.data? (Use class() function)
```{r}
class(abalone.data)
```

\newpage
##### Data Clean Up
*Add Column Names to Data Frame*
```{r}
colnames(abalone.data)[1] <- "sex"
colnames(abalone.data)[2] <- "length"
colnames(abalone.data)[3] <- "diameter"
colnames(abalone.data)[4] <- "height"
colnames(abalone.data)[5] <- "whole_weight"
colnames(abalone.data)[6] <- "shucked_weight"
colnames(abalone.data)[7] <- "viscera_weight"
colnames(abalone.data)[8] <- "shell_weight"
colnames(abalone.data)[9] <- "rings"
```

*New Output*
```{r}
head(abalone.data)
```

\newpage
### 2) What is the datatype of variable diameter? (Use typeof() function)
```{r}
typeof(abalone.data$diameter)
```

\newpage
### 3) Use the function summary() to find basic description of diameter and age (rings) of abalone.
```{r}
summary(abalone.data$diameter)
```
```{r}
summary(abalone.data$rings)
```

\newpage
### 4) Use the function mean() to find the mean of diameter of female abalone.
```{r}
female.abalone <- dplyr::filter(abalone.data,sex=='F')
mean(female.abalone$diameter)
```
https://rstudio-pubs-static.s3.amazonaws.com/408480_f4bdfd9620c84a9598e512f1a59e66f5.html

\newpage
### 5) Assume Y: diameter and X: rings. What is〖(X^' X)〗^(-1) X^' Y ? 
```{r}
x <- abalone.data$rings
y <- abalone.data$diameter

transposed <- t(x) %*% x
inversed <- solve(transposed)
part2 <- t(x) %*% y
answer <- inversed %*%part2
answer

```


\newpage
### 6) Can you think of an interesting question about Abalone? 
I had no real goals for Abalone, but instead chose to take this opportunity to  explore the different packages and capabilities RStudio provides. I ran histograms, side-by-side-by-side box plots, linear regression models, scatter plots of those models, sensitivity analysis of a model, and statistical tests of covariance, correlation, collinearity, and fit of analysis for the regression model. 

The biggest improvement I would make to this section is the output. Instead of having 4 histograms all follwing each other taking up a half a page or even more, I would manipulate the size of the output so all 4 histograms would be on the same page in a 2x2 table. 

*I need to make the data and graphs more readable.* 

The first step in any analysis was to continue the separation of the abalone by sex.  Since the females were previously partitioned, only the males and infants are required 

```{r}
male.abalone=dplyr::filter(abalone.data,sex=='M')
indet.abalone=dplyr::filter(abalone.data,sex=='I')
```

Now we can run a plots and tests for the data as a whole.
Below is a histogram of the entire data with a focus on the rings, length, weight, and diameter by total data set, male, female, and indeterminate genders.
```{r}
hist(abalone.data$rings)
hist(male.abalone$rings)
hist(female.abalone$rings)
hist(indet.abalone$rings)
```

\newpage
Here are the histograms of the data focused on length by gender
```{r}
hist(abalone.data$length)
hist(male.abalone$length)
hist(female.abalone$length)
hist(indet.abalone$length)
```


\newpage
Here are the histograms of the data focused on whole weight by gender
```{r}
hist(abalone.data$whole_weight)
hist(male.abalone$whole_weight)
hist(female.abalone$whole_weight)
hist(indet.abalone$whole_weight)
```


\newpage
Here are the histograms of the data focused on diameter by gender
```{r}
hist(abalone.data$diameter)
hist(male.abalone$diameter)
hist(female.abalone$diameter)
hist(indet.abalone$diameter)
```


\newpage
Here are the histograms of the data focused on height by gender
```{r}
hist(abalone.data$height)
hist(male.abalone$height)
hist(female.abalone$height)
hist(indet.abalone$height)
```

\newpage
After running through some new statistical tests, I went back to last weeks assignment and created a side-by-side boxplot.
As you can see, there is no real difference in ages based upon gender. 
This is to be expected as Abalone do not have a natural cannibalistic relationship like the praying mantis or black widow. 
One thing I would like to expound upon further is the utilization of ggplot2 to increase the color schema for readability.

```{r}
Male.Rings = abalone.data$rings[which(abalone.data$sex=='M')] 
Female.Rings = abalone.data$rings[which(abalone.data$sex=='F')]
Indet.Rings = abalone.data$rings[which(abalone.data$sex=='I')]
my.bp <- boxplot(Male.Rings, Female.Rings, Indet.Rings,
                 main = "Number of Rings by Gender",
                 ylab = "Ages",
                 names = c("Male", "Female", "Indeterminate"))

```

```{r}
Male.Length = abalone.data$length[which(abalone.data$sex=='M')] 
Female.Length = abalone.data$length[which(abalone.data$sex=='F')]
Indet.Length = abalone.data$length[which(abalone.data$sex=='I')]
my.bp <- boxplot(Male.Length, Female.Length, Indet.Length,
                 main = "Length by Gender",
                 ylab = "Ages",
                 names = c("Male", "Female", "Indeterminate"))

```

```{r}
Male.WWeight = abalone.data$whole_weight[which(abalone.data$sex=='M')] 
Female.WWeight = abalone.data$whole_weight[which(abalone.data$sex=='F')]
Indet.WWeight = abalone.data$whole_weight[which(abalone.data$sex=='I')]
my.bp <- boxplot(Male.WWeight, Female.WWeight, Indet.WWeight,
                 main = "Whole Weight by Gender",
                 ylab = "Ages",
                 names = c("Male", "Female", "Indeterminate"))

```

```{r}
Male.Diameter = male.abalone$diameter
Female.Diameter = female.abalone$diameter
Indet.Diameter = indet.abalone$diameter
my.bp <- boxplot(Male.Diameter, Female.Diameter, Indet.Diameter,
                 main = "Diameter by Gender",
                 ylab = "Ages",
                 names = c("Male", "Female", "Indeterminate"))

```

```{r}
Male.Height = abalone.data$height[which(abalone.data$sex=='M')] 
Female.Height = abalone.data$height[which(abalone.data$sex=='F')]
Indet.Height = abalone.data$height[which(abalone.data$sex=='I')]
my.bp <- boxplot(Male.Height, Female.Height, Indet.Height,
                 main = "Height by Gender",
                 ylab = "Height",
                 names = c("Male", "Female", "Indeterminate"))

```

\newpage
Some basic correlation, covariance, and standar deviation calculations
```{r}
corLengthAge <- cor.test(abalone.data$length, abalone.data$rings)
corWeightAge <- cor.test(abalone.data$whole_weight, abalone.data$rings) 
corHeightAge <- cor.test(abalone.data$height, abalone.data$rings)

covLengthAge <- cov(abalone.data$length, abalone.data$rings)
covWeighthAge <- cov(abalone.data$whole_weight, abalone.data$rings)
covHeighthAge <- cov(abalone.data$height, abalone.data$rings)

corLengthAge
corWeightAge
corHeightAge

covLengthAge
covWeighthAge
covHeighthAge

sd(abalone.data$rings)
```
\newpage
Smooth scatter plot for rings vs whole weight, diameter.
Smooth scatter plot for whole weight and length vs diameter.
Smooth scatter plot for length vs height
```{r}
scatter.smooth(x=abalone.data$rings, y=abalone.data$whole_weight,
               xlab = "Rings",
               ylab = "Whole Weight"
)

```

```{r}
scatter.smooth(x=abalone.data$rings, y=abalone.data$diameter,
               xlab = "Rings",
               ylab = "Diameter"
)

```

```{r}
scatter.smooth(x=abalone.data$whole_weight, y=abalone.data$diameter,
               xlab = "Whole Weight",
               ylab = "Diameter"
)

```

```{r}
scatter.smooth(x=abalone.data$length, y=abalone.data$diameter,
               xlab = "Length",
               ylab = "Diameter"
)

```

```{r}
scatter.smooth(x=abalone.data$length, y=abalone.data$height,
               xlab = "Length",
               ylab = "Height"
)

```

```{r}
scatter.smooth(x=abalone.data$height, y=abalone.data$diameter,
               xlab = "Length",
               ylab = "Height"
)

```

\newpage
Additional least squares estimation via matrix math.
Entirely focused on whole weight vs shucked weight, viscera weight, length, diameter, and height. 
```{r}
ww.matrix <- abalone.data$whole_weight
sw.matrix <- abalone.data$shucked_weight
v.matrix <- abalone.data$viscera_weight
l.matrix <- abalone.data$length
d.matrix <- abalone.data$diameter
h.matrix <- abalone.data$height
head(abalone.data)

t.ww <- t(ww.matrix) %*% ww.matrix
inversed.ww <- ginv(ww.matrix)
part2.sw <- t(ww.matrix) %*% sw.matrix
answer.sw <- inversed %*%part2.sw
answer.sw
```

```{r}
part2.v <- t(ww.matrix) %*% v.matrix
answer.v <- inversed %*%part2.v
answer.v
```

```{r}
part2.l <- t(ww.matrix) %*% l.matrix
answer.l <- inversed %*%part2.l
answer.l
```

```{r}
part2.d <- t(ww.matrix) %*% d.matrix
answer.d <- inversed %*%part2.d
answer.d
```

```{r}
part2.h <- t(ww.matrix) %*% h.matrix
answer.h <- inversed %*%part2.h
answer.h
```

\newpage
Linear Regression models
  Whole weight vs Rings and Diameter
  Viscera Weight vs Whole Weight and Shell Weight
  Whole Weight vs Diameter and Height and Length
```{r}
lm.ring <- lm(abalone.data$rings ~ abalone.data$diameter)
lm.ring
```

```{r}
lm.vweight <- lm(abalone.data$viscera_weight ~ abalone.data$shell_weight + abalone.data$whole_weight)
lm.vweight
```

```{r}
lm.wweight <- lm(abalone.data$whole_weight ~ 
                   abalone.data$diameter + 
                   abalone.data$height + 
                   abalone.data$length)
lm.wweight
```

\newpage
Testing collinearity using VIF function
```{r}
car::vif(lm.wweight)
```

\newpage
AOV tests regression fit
```{r}
aov(lm.ring)
```

```{r}
aov(lm.vweight)
```

```{r}
aov(lm.wweight)
```

\newpage
Summaries of the three regression tests
```{r}
summary(lm.ring)
```

```{r}
summary(lm.vweight)
```

```{r}
summary(lm.wweight)
```

\newpage
Plot of Rings vs Diameter
```{r}
plot1 <- plot(abalone.data$rings ~ abalone.data$diameter, data = abalone.data)
```

Plot of Viscera Weight vs Shell Weight
```{r}
plot(abalone.data$viscera_weight ~ abalone.data$shell_weight, data = abalone.data)
```

Plot of Whole Weight vs Diameter
```{r}
plot(abalone.data$whole_weight ~ abalone.data$diameter, data = abalone.data)
```

Sensitivity analysis
```{r}
wweight.sensitivity <- sensemakr(model = lm.wweight, 
                                treatment = "abalone.data$height",
                                benchmark_covariates = "abalone.data$diameter",
                                kd = 1:3,
                                ky = 1:3, 
                                q = 1,
                                alpha = 0.05, 
                                reduce = TRUE)
wweight.sensitivity
```

\newpage
https://cran.r-project.org/web/packages/sensemakr/vignettes/sensemakr.html

http://faculty.cas.usf.edu/mbrannick/regression/regma.htm

https://escholarship.org/content/qt2k89z4z7/qt2k89z4z7_noSplash_79a83593875d17076c19b260e6c60fd4.pdf

http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r#install-and-load-required-r-packages

https://rstudio-pubs-static.s3.amazonaws.com/408480_f4bdfd9620c84a9598e512f1a59e66f5.html

http://www.sthda.com/english/articles/39-regression-model-diagnostics/160-multicollinearity-essentials-and-vif-in-r/

https://github.com/nishitpatel01/predicting-age-of-abalone-using-regression
