---
title: "M2A Assignment"
author: "Andrew Estes"
date: "8/22/2021"
output: word_document
---

##Install Packages
```{r}
library(ggpubr)
library(car)
library(caret)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(MASS)
```

##Importing and View Data
```{r}
abalone.url <-
  "https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data"
abalone.data <- read.csv(abalone.url, header=FALSE) #data does not have variable names
abalone.data

summary(abalone.data)
head(abalone.data)
```

##1) What is the class of the data abalone.data? (Use class() function)
```{r}
class(abalone.data)
```


####Add Column Names to Data Frame
```{r}
colnames(abalone.data)[1] <- "sex"
colnames(abalone.data)[2] <- "length"
colnames(abalone.data)[3] <- "diameter"
colnames(abalone.data)[4] <- "height"
colnames(abalone.data)[5] <- "whole_weight"
colnames(abalone.data)[6] <- "shucked_weight"
colnames(abalone.data)[7] <- "viscera_weight"
colnames(abalone.data)[8] <- "shell_weight"
colnames(abalone.data)[9] <- "rings"

summary(abalone.data)
head(abalone.data)
```

##2) What is the datatype of variable diameter? (Use typeof() function)
```{r}
typeof(abalone.data$diameter)
```

##3) Use the function summary() to find basic description of diameter and age (rings) of abalone.
```{r}
summary(abalone.data$diameter)
```
```{r}
summary(abalone.data$rings)
```

##4) Use the function mean() to find the mean of diameter of female abalone.
```{r}
female.abalone=dplyr::filter(abalone.data,sex=='F')
summary(female.abalone)
mean(female.abalone$diameter)
```
https://rstudio-pubs-static.s3.amazonaws.com/408480_f4bdfd9620c84a9598e512f1a59e66f5.html


##5) Assume Y: diameter and X: rings. What is〖(X^' X)〗^(-1) X^' Y ? 
```{r}
X <- list(abalone.data$rings)
Y <- list(abalone.data$diameter)
Z <- c(X,Y)

abalone.matrix <- matrix(unlist(Z), ncol=2)
colnames(abalone.matrix) <- c("Rings", "Diameter")
abalone.matrix

head(abalone.matrix, 2)

t(abalone.matrix) #transposed
ginv(abalone.matrix) #inverse

```

###6) Can you think of an interesting question about Abalone? 
First step is to create the male abalone data.
```{r}
male.abalone=dplyr::filter(abalone.data,sex=='M')
```

Now we can run a plots and tests for the data as a whole
```{r}
hist(abalone.data$rings)

corLengthAge <- cor.test(abalone.data$length, abalone.data$rings)
corWeightAge <- cor.test(abalone.data$whole_weight, abalone.data$rings) 
corHeightAge <- cor.test(abalone.data$height, abalone.data$rings)

covLengthAge <- cov(abalone.data$length, abalone.data$rings)
covWeighthAge <- cov(abalone.data$whole_weight, abalone.data$rings)
covHeighthAge <- cov(abalone.data$height, abalone.data$rings)

sd(abalone.data$rings)

scatter.smooth(x=abalone.data$rings, y=abalone.data$whole_weight,
               xlab = "Rings",
               ylab = "Whole Weight"
)

```

Now we can break it down by gender
```{r}
Male = abalone.data$rings[which(abalone.data$sex=='M')] 
Female = abalone.data$rings[which(abalone.data$sex=='F')]
my.bp <- boxplot(Male,Female,
                 main = "Number of Rings by Gender",
                 ylab = "Ages",
                 names = c("Male", "Female"))

```


Here is a predictive model that I copy+pasted and do not fully understand
```{r}
malemodel <- lm(rings~length+diameter+height+whole_weight+viscera_weight+shell_weight,data=male.abalone)
femalemodel <- lm(rings~length+diameter+height+whole_weight+viscera_weight+shell_weight,data=female.abalone)

x <- lm(rings~length,data=male.abalone)
y <- lm(rings~length,data=female.abalone)


set.seed(123)
age.samples <- abalone.data$rings %>%
  createDataPartition(p = 0.8, list = FALSE)
train.data <- abalone.data[age.samples, ]
test.data <- abalone.data[-age.samples, ]

model1 <- lm(rings ~., data = train.data)

predictions <- model1 %>% predict(test.data)

data.frame(
  RMSE = RMSE(predictions, test.data$rings),
  R2 = R2(predictions, test.data$rings)
)

car::vif(model1)

model2 <- lm(rings ~. -whole_weight -viscera_weight -diameter -length, data = train.data)
#reduced number of variables using correlation/covariance data from the top part of this question

predictions <- model2 %>% predict(test.data)

data.frame(
  RMSE = RMSE(predictions, test.data$rings),
  R2 = R2(predictions, test.data$rings)
)

car::vif(model2)
```
http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r#install-and-load-required-r-packages
https://rstudio-pubs-static.s3.amazonaws.com/408480_f4bdfd9620c84a9598e512f1a59e66f5.html
http://www.sthda.com/english/articles/39-regression-model-diagnostics/160-multicollinearity-essentials-and-vif-in-r/
https://github.com/nishitpatel01/predicting-age-of-abalone-using-regression







